Manuscript draft

### 2019-07-17 manuscript/manuscript draft.Rmd
updated manuscript draft

*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect. Site altitude was also included as a fixed effect. Random effect was stand nested within area. Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest. These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone. Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.

For each climate variable, two models were specified: one with the climate variable as a linear term, and one with the climate variable as a polynomial term. AIC values were calculated for the models and used to select the best model. R-squared values were calculated for the best model. In addition, slope values were calculated for the linear models.

@@

>>>>

*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect. Site altitude was also included as a fixed effect. Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas. To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area. Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest. These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone. Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.

For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term. AIC values were calculated for the models and used to select the best model. R-squared values were calculated for the best model. In addition, slope coefficients were calculated for the linear models.

@@

To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified. Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 8 climate variables. For each possible pairing of climate variables, three models were specified: a null model; a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect. As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect. AIC values were calculated for the models and used to select the best model. R-squared values were calculated for the best model. Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.
